---
title: "Untitled"
author: "Zachary M. Smith"
date: "December 5, 2018"
output: html_document
---



## Calculate Metric Sensitivity

```{r}
reg.df <- metrics.long %>% 
  filter(lg_reg_v2 %in% c("Hudson Valley", "Southern Tier"))

```


### State-Wide Metric Sensitivity

Barbour et al. (1996) created a visual method for evaluating metric response to a defined disturbance gradient.

* 0 = The medians of both the reference and degraded conditions overlap the interquartile range of one another.
* 1 = One median (either the reference or degraded) overlaps the interquartile range of the other.
* 2 = The interquartile ranges of the reference and degraded conditions overlap but neither conditions median overlaps with the interquartile range of the other category.
* 3 = The interquartile ranges of the reference and degraded conditions do not overlap.

Discrimination Efficiency (DE) is a measure of the percentage of degraded samples correctly identified below the reference distributions 25th percentile for metrics that decrease with degradation or the percentage of degraded samples correctly identified above the reference distributions 75th percentile for metrics that increase with disturbance.

Balanced Discrimination Efficiency (BDE) attempts to update the DE method by throwing more computation power at the problem. BDE tests multiple thresholds for separating the reference and degraded distributions in an attempt to find the best separation point.

Use the `sensitivity()` function to calculate three types of metric sensitivity measures: 1) Barbour et al. (1996) method, 2) Discrimination Efficiency (DE), and 3) Balanced Discrimination Efficiency (BDE). 
```{r}
sw.sens.df <- reg.df %>% 
  sensitivity(
    metric.col = metric,
    value.col = value,
    condition.col = site_condition,
    ref.cond = "ref",
    deg.cond = "deg")

DT::datatable(sw.sens.df, options = list(scrollX = TRUE))
```

#### Plotting

Plot Barbour et al. (1996) values greater than or equal to 2.
```{r, fig.width=10, fig.height=1.75}
metrics.vec <- c("pct_et",
                 "pct_rich_cote_family",
                 "rich_toe_family",
                 "pct_filterer",
                 "pct_rich_odonata",
                 "shannon_genus"
                 )

sw.sens.df %>% 
  filter(metric %in% metrics.vec) %>%
  # filter(barbour == 3) %>%
  # filter(de >= 70) %>%
  inner_join(reg.df, by = "metric") %>% 
  mutate(metric = factor(metric, levels = unique(metric))) %>% 
  ggplot(aes(site_condition, value)) +
  geom_boxplot() +
  # geom_jitter() +
  facet_wrap(~metric, ncol = 6, scales = "free")
  # Use line below instead for plotting metrics for each region side-by-side
  # facet_wrap(~metric + lg_reg_v2, ncol = 3, scales = "free")


```


```{r}
sub.df <- reg.df %>% 
  filter(metric %in% metrics.vec) %>% 
  group_by(metric) %>% 
  mutate(quant_5 = quantile(value, 0.05),
         quant_95 = quantile(value, 0.95),
         min = min(value),
         max = max(value),
         score = (value - quant_5) / (quant_95 - quant_5) * 100,
         score = case_when(
           score > 100 ~ 100,
           score < 0 ~ 0,
           TRUE ~ score
         )) %>% 
  select(-min, -max, -value)

index.df <- sub.df %>% 
  # spread(metric, score) %>% 
  group_by(bas_loc_rm, site_condition, lg_reg_v2) %>% 
  summarise(index_score = mean(score)) %>% 
  ungroup()
```

```{r}
theme_set(theme_grey(base_size = 20)) 
index.df %>% 
  ggplot(aes(site_condition, index_score)) +
  geom_boxplot() 


```

```{r}
index.df %>% 
  # filter(site_condition == "ref") %>% 
  ggplot(aes(lg_reg_v2, index_score, fill = site_condition)) +
  geom_boxplot()
```



```{r}
selected.df <- sub.df %>% 
  spread(metric, score) %>% 
  left_join(condition.df, by = c("bas_loc_rm", "site_condition", "lg_region")) %>% 
  select(site_condition, metrics.vec) %>% 
  mutate(site_condition = if_else(site_condition == "ref", 1, 0))


logistic <- glm(site_condition ~ ., data = selected.df, family = "binomial")

selected.df$predicted <- predict(logistic, newdata = selected.df, type="response")

coef.df <- summary(logistic)$coefficients %>% 
  as_tibble() %>% 
  mutate(metric = rownames(summary(logistic)$coefficients)) %>% 
  filter(`Pr(>|z|)` < 0.1,
         metric != "(Intercept)") %>% 
  pull(metric)


selected.df <- sub.df  %>% 
  spread(metric, score) %>% 
  left_join(condition.df, by = c("bas_loc_rm", "site_condition", "lg_region")) %>% 
  select(site_condition, coef.df) %>% 
  mutate(site_condition = if_else(site_condition == "ref", 1, 0))


logistic <- glm(site_condition ~ ., data = selected.df, family = "binomial")

selected.df$predicted <- predict(logistic, newdata = selected.df, type="response")

```

```{r, fig.width=10, fig.height=4}
sw.sens.df %>% 
  filter(metric %in% coef.df) %>% 
  inner_join(metrics.long, by = "metric") %>% 
  mutate(metric = factor(metric, levels = unique(metric))) %>% 
  ggplot(aes(site_condition, value)) +
  geom_boxplot() +
  facet_wrap(~metric, ncol = 3, scales = "free")
```

```{r, fig.width=10, fig.height=4}
sw.sens.df %>% 
  filter(metric %in% coef.df) %>% 
  inner_join(metrics.long, by = "metric") %>% 
  mutate(metric = factor(metric, levels = unique(metric))) %>% 
  ggplot(aes(site_condition, value, fill = lg_region)) +
  geom_boxplot() +
  facet_wrap(~metric, ncol = 3, scales = "free")
```

```{r}
tp <- nrow(selected.df[selected.df$site_condition == 1 & selected.df$predicted >= 0.5, ])
ref.n <- nrow(selected.df[selected.df$site_condition == 1, ])
tp.pct <- tp / ref.n * 100

fp <- nrow(selected.df[selected.df$site_condition == 0 & selected.df$predicted < 0.5, ])
test.n <- nrow(selected.df[selected.df$site_condition == 0, ])
fp.pct <- fp / test.n * 100

(tp.pct + fp.pct) / 2
```

```{r}
ggplot(selected.df, aes(x=predicted, y=site_condition)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) 
```

```{r}
ggplot(selected.df, aes(as.factor(site_condition), predicted)) +
  geom_boxplot() +
  geom_jitter()
```