---
title: "Abiotic_corr_matrices"
author: "Gavin Lemley"
date: "January 24, 2019"
output: html_document
---

Find the R-project root directory to simplify file path designations.
```{r}
root.dir <- rprojroot::find_root("lgss.Rproj")
```

Load in data table
```{r}

abiotic.df <- read.csv(file.path(root.dir, "data/Compile_env_data", "LGSS_2015_env_numerical_ALL.csv"))

abiotic.df$BAS_LOC_RM <- NULL 

#Subset data by parameter (column number) for more legible plots
abiotic.dfsmall.1 <- abiotic.df[,c(1:6)]
abiotic.dfsmall.2 <- abiotic.df[,c(7:12)]
abiotic.habitat <- abiotic.df[,c(13:25)]
abiotic.dfsmall.4 <- abiotic.df[,c(26:33)]

abiotic.nohabitat <- abiotic.df[ -c(13:25) ]

```

Plot correlation matrix
```{r}
pairs(abiotic.df[,1:6], pch = 19)
```

Using package psych 
```{r}
psych::pairs.panels(abiotic.dfsmall.1[,-5], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
psych::pairs.panels(abiotic.dfsmall.2[,-5], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
psych::pairs.panels(abiotic.habitat[,-5], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
psych::pairs.panels(abiotic.dfsmall.4[,-5], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```

Functions sent from Brian
```{r}
#function from "pairs function help file" to calculate correlation between variables, also shown on pg 535 in Zuur 2009
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}

#function from "pairs function help" file to produce histogram of each variable, also shown on pg 535 in Zuur 2009
panel.hist <- function(x, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5) )
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}

#From page 535 of Zuur 2009 book, matrix of all covariates (and the response variable) to look for colinearity
  #covariates with high corelations (e.g. >0.6) should potentially be excluded
# AllVars <- cbind(BAP$FINAL_SCORE, BAP$Depth, BAP$Temp, BAP$SpC, 
#                  BAP$pH, BAP$DO, BAP$Secchi, BAP$perTOC, BAP$Phi, BAP$perDreissena,
#                  BAP$chinumsurv, BAP$chiavewieght,BAP$chiavebiom,BAP$hynumsurv,BAP$hyavewieght,BAP$hyavebiom)
# colnames(AllVars) <- c("BAP", "Depth", "Temp", "SpC", "pH", "DO", "Secchi", "perTOC", "Phi", "%Dreissena",
#                        "chi#","chiW","chiB","hya#","hyaW","hyaB")
pairs(abiotic.dfsmall.1, upper.panel = panel.cor, diag.panel = panel.hist)

```

Using corrplot package
```{r,fig.width=20, fig.height=20}
M<-cor(abiotic.df, use = "complete.obs" )
head(round(M,2))

# See http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram for different graph styles

corrplot::corrplot(M, method="circle")

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot::corrplot(M, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         # p.mat = p.mat, sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )



```

Corrplot for corr values limited to >0.5 
```{r,fig.width=20, fig.height=20}

M.df <- as.data.frame(M)
M.df <- apply(M.df,2,abs)
col.i <- 1
keep.vec <- sapply(1:ncol(M.df), function(col.i){
  sub.vec <- M.df[, col.i]
  any(sub.vec > 0.5 & sub.vec < 1)
})
smallmatrix.df <- M[keep.vec,keep.vec]


col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot::corrplot(smallmatrix.df, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         # p.mat = p.mat, sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )
```

Corrplot for corr values limited to >0.7 
```{r,fig.width=10, fig.height=10}

M.df <- as.data.frame(M)
M.df <- apply(M.df,2,abs)
col.i <- 1
keep.vec <- sapply(1:ncol(M.df), function(col.i){
  sub.vec <- M.df[, col.i]
  any(sub.vec > 0.7 & sub.vec < 1)
})
smallmatrix.df <- M[keep.vec,keep.vec]


col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot::corrplot(smallmatrix.df, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         # p.mat = p.mat, sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )
```


### Below chunks look at DF without habitat variables 

Using corrplot package
```{r,fig.width=20, fig.height=20}
M<-cor(abiotic.nohabitat, use = "complete.obs" )
head(round(M,2))

# See http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram for different graph styles

corrplot::corrplot(M, method="circle")

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot::corrplot(M, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         # p.mat = p.mat, sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )



```

Corrplot for corr values limited to >0.5 
```{r,fig.width=10, fig.height=10}

M.df <- as.data.frame(M)
M.df <- apply(M.df,2,abs)
col.i <- 1
keep.vec <- sapply(1:ncol(M.df), function(col.i){
  sub.vec <- M.df[, col.i]
  any(sub.vec > 0.5 & sub.vec < 1)
})
smallmatrix.df <- M[keep.vec,keep.vec]


col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot::corrplot(smallmatrix.df, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         # p.mat = p.mat, sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )
```

Corrplot for corr values limited to >0.7 
```{r,fig.width=10, fig.height=10}

M.df <- as.data.frame(M)
M.df <- apply(M.df,2,abs)
col.i <- 1
keep.vec <- sapply(1:ncol(M.df), function(col.i){
  sub.vec <- M.df[, col.i]
  any(sub.vec > 0.7 & sub.vec < 1)
})
smallmatrix.df <- M[keep.vec,keep.vec]


col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot::corrplot(smallmatrix.df, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         # p.mat = p.mat, sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )
```


