---
title: "compile_2015-2019_abiotic"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)

root.dir <- rprojroot::find_root("lgss.Rproj")

fielddata.2015.path <- file.path(root.dir, "data", "compile_2015-2019", "field_data", "2015_LGSS_data_MASTER_oct2020_siteIDcorrections.xlsx")
fielddata.2016.2017.path <- file.path(root.dir, "data", "compile_2015-2019", "field_data", "LGSS_16-17_MASTER_v2_noNA.xlsx")

itstable.path <- file.path(root.dir, "data", "compile_2015-2019", "ITS_tables_20201030")

teams.datadir <- "C:/Users/gmlemley/New York State Office of Information Technology Services/SMAS - Low Gradient IBI/data"

```

Field tables: Read in best versions and join
```{r}

# Full StreamStats output tables are in the same dir as in case more variables are needed.

field.LU.2015 <- read_excel(fielddata.2015.path, sheet = "FieldData") %>% 
  select(-BAS_LOC_RM...39, -BAS_LOC_RM_gis) %>% 
  janitor::clean_names() %>% 
  rename_all(tolower) %>% 
  select(-c(name, direct, descript, lat_final, long_final, cond_test_250, land_use_pct_natural_wetland)) %>% 
  mutate(basin = as.numeric(basin)) %>% 
  rename(site_id = bas_loc_rm_1,
         sample_date = date_field_data,
         arrivetime = time,
         lgs_habitat_bank = bank,
         lgs_habitat_macrophyte = macrophyte,
         lgs_habitat_snag = snags,
         lgs_habitat_substrate = substrate,
         ref_field_yn = reference_fieldsheet_question,
         general_comment = notes,
         macrophyte_pct = percent_macrophyte,
         cur_spd = current,
         spcond = conduct,
         ph = p_h,
         do_conc = do,
         do_pct_sat = percent_sat
  ) %>% 
  mutate(arrivetime = format(arrivetime, format="%H:%M"))

field.LU.2016.2017 <- read_excel(fielddata.2016.2017.path, sheet = "LG data + sites, LU, SteamStats") %>%  
  select(1:62,84) %>% 
  janitor::clean_names() %>% 
  rename_all(tolower) %>% 
  setNames(., sub("_x", "", names(.))) %>% 
  select(-c(location, rivmile, station, org_for_toxics, project, samp, secchi, lu_pct_nat_wetlnd)) %>% 
  rename(site_id = bas_loc_rm,
         sample_date = coll_date,
         embeddedness = embedd,
         do_conc = do,
         do_pct_sat = percent_sat,
         spcond = cond,
         algae_suspended = algae_s,
         algae_filamentous = algae_fil,
         algae_diaoms_presen = algae_d,
         thickness_diatoms_present = thickness_d,
         field_water_qlty = field_wqa,
         lgs_habitat_bank = lgs_hab_bnk,
         lgs_habitat_macrophyte = lgs_hab_mac,
         lgs_habitat_snag = lgs_hab_sng,
         lgs_habitat_substrate = lgs_hab_sub,
         dominant_substrate = dom_substrate,
         water_clarity_driver = wtrclr_driver,
         land_use = landuse,
         general_comment = notes
  ) %>% 
  mutate(arrivetime = format(arrivetime, format="%H:%M"))

# Bind 2015 and 2016+2017 data together
field.LU.bind <- bind_rows(field.LU.2015, field.LU.2016.2017) 
  # select(sort(current_vars()))

field.LU.bind <- field.LU.bind %>% 
  select(-c(algae_suspended, algae_filamentous, "ephem", "plecop", "trich", "coleop", "megalop", "odonata", "chir", "simul", "decapod", "gammar", "mollusk", "oligo", "others"))


# Field names for referece (binding)

# > names(field.LU.2015)
#  [1] "bas_loc_rm_1"                  "basin"                         "name"                          "direct"                        "descript"                     
#  [6] "lat_final"                     "long_final"                    "date_field_data"               "time"                          "bank"                         
# [11] "macrophyte"                    "crew"                          "notes"                         "photo"                         "photo_2"                      
# [16] "snags"                         "substrate"                     "reference_fieldsheet_question" "percent_macrophyte"            "depth"                        
# [21] "bottom_type"                   "width"                         "dominant_substrate"            "current"                       "riffle_present"               
# [26] "canopy"                        "land_use"                      "algae_filamentous"             "algae_suspended"               "temp"                         
# [31] "p_h"                           "do"                            "conduct"                       "percent_sat"                   "chl_a"                        
# [36] "water_clarity_driver"          "land_use_pct_natural_wetland"  "site_condition_v1"             "site_condition_v2"             "cond_test_250"                
# [41] "site_condition_notes" 
# 
# > names(field.LU.2016.2017)
#  [1] "bas_loc_rm"        "basin"             "coll_date"         "crew"              "arrivetime"        "samp"              "landuse"           "depth"            
#  [9] "width"             "cur_spd"           "canopy"            "embedd"            "rock"              "rubble"            "gravel"            "sand"             
# [17] "silt"              "clay"              "temp"              "cond"              "do"                "ph"                "percent_sat"       "salinity"         
# [25] "secchi"            "algae_s"           "algae_fil"         "algae_d"           "thickness_d"       "macrophyte_pct"    "ephem"             "plecop"           
# [33] "trich"             "coleop"            "megalop"           "odonata"           "chir"              "simul"             "decapod"           "gammar"           
# [41] "mollusk"           "oligo"             "others"            "field_wqa"         "habitat"           "org_for_toxics"    "microtox"          "lgs_hab_bnk"      
# [49] "lgs_hab_mac"       "lgs_hab_sng"       "lgs_hab_sub"       "bottom_type"       "dom_substrate"     "wtrclr_driver"     "pht_id_up"         "pht_id_dwn"       
# [57] "notes"             "project"           "hab_present"       "lu_pct_nat_wetlnd"       
# #          
# > names(db.SEI)
#  [1] "instance_id"               "row_index"                 "event_smas_id"             "event_smas_history_id"     "basin"                     "location"                 
#  [7] "rivermile"                 "event_smas_sample_date"    "smas_arrivetime"           "crew"                      "site_type_dept_intrst"     "site_type_extrnl_sugg"    
# [13] "site_type_lng_trm_trnd"    "site_type_mltplt"          "site_type_probabilistc"    "site_type_ras"             "site_type_reference"       "site_type_rmn"            
# [19] "site_type_spcl_stdy"       "site_type_unassd_pwl"      "event_baseline"            "event_extent"              "disch_collect"             "total_disch"              
# [25] "biosample_collect"         "biosample_type"            "water_chem_collect"        "chem_qaqc"                 "chem_qaqc_type"            "chem_samp_loc"            
# [31] "chem_samp_equp"            "microtox"                  "org_for_toxics"            "depth"                     "depth_qlfr"                "width"                    
# [37] "width_qlfr"                "speed_of_stream_current"   "spd_of_strm_current_qlfr"  "secchi"                    "secchi_qlfr"               "algae_diaoms_presen"      
# [43] "thickness_diatoms_present" "algae_suspended"           "algae_filamentous"         "mfdh_macrophytes"          "canopy"                    "embeddedness"             
# [49] "embeddedness_qlfr"         "rock"                      "rubble"                    "gravel"                    "sand"                      "silt"                     
# [55] "clay"                      "land_use"                  "general_comment"           "chem_result_rcvd"          "chem_result_rcvd_note"    
# 
# > names(db.insitu)
#  [1] "row_index"                  "instance_id"                "event_smas_id"              "event_smas_history_id"      "basin"                      "location"  
#  [7] "rivermile"                  "event_smas_sample_date"     "smas_arrivetime"            "ysi_meter_id"               "chem_parameter_pcode_valid" "result"       
#  [13] "validator_qual"             "validator_qual_reason"      "validation_date"            "qlfr_note"                 
# 
# > names(db.macrofield)
#  [1] "row_index"                "instance_id"              "event_smas_id"            "event_smas_history_id"    "basin"                    "location"                
#  [7] "rivermile"                "event_smas_sample_date"   "biosample_collect_method" "biosamp_col_method_note"  "ephemeroptera"            "plecoptera"              
# [13] "trichoptera"              "coleoptera"               "megaloptera"              "odonata"                  "chironomidae"             "simuliidae"              
# [19] "decapods"                 "gammaridae"               "mollusca"                 "oligochaeta"              "tipulidae"                "others"                  
# [25] "other"                    "field_water_qlty"         "habitat"                  "lgs_habitat_bank"         "lgs_habitat_macrophyte"   "lgs_habitat_snag"        
# [31] "lgs_habitat_substrate"    "bottom_type"              "dominant_substrate"       "water_clarity_driver"     "macro_comment" 


rm(field.LU.2015, field.LU.2016.2017)

```

Habitat tables: Read in best versions and join
```{r}
habitat.2015 <- read_excel(fielddata.2015.path, sheet = "Habitat") %>% 
  janitor::clean_names() %>% 
  rename_all(tolower) %>% 
  # mutate(date_habitat = ifelse(bas_loc_rm %in% "01-BLKE-0.1", date_habitat+1, date_habitat)) %>%
  rename(site_id = bas_loc_rm,
         sample_date = date_habitat,
         epifaunal_cover = x1_epifaunal_substrate_available_cover,
         embeddedness_pooling = x2_pool_substrate_characterization,
         velocity_depth_regime = x3_pool_variability,
         sediment_deposition = x4_sediment_deposition,
         flow_status = x5_channel_flow_status,
         channel_alteration = x6_channel_alteration,
         riffle_bend_frequency = x7_channel_sinuosity,
         left_bank_stability = x8_lb_stability,
         right_bank_stability = x8_rb_stability,
         left_bank_veg = x9_lb_vegetative_protection,
         right_bank_veg = x9_rb_vegetative_protection,
         left_bank_veg_zone = x10_lb_riparian_zone_width,
         right_bank_veg_zone = x10_rb_riparian_zone_width,
         habitat_comment = comments)

# > names(db.habitat)
#  [1] "row_index"              "event_smas_history_id"  "basin"                  "location"               "rivermile"              "event_smas_sample_date"
#  [7] "gradient"               "epifaunal_cover"        "embeddedness_pooling"   "velocity_depth_regime"  "sediment_deposition"    "flow_status"           
# [13] "channel_alteration"     "riffle_bend_frequency"  "left_bank_stability"    "right_bank_stability"   "left_bank_veg"          "right_bank_veg"        
# [19] "left_bank_veg_zone"     "right_bank_veg_zone"    "habitat_comment"        "instance_id"            "event_smas_id"  

# > names(habitat.2015)
#  [1] "bas_loc_rm"                             "date_habitat"                           "x1_epifaunal_substrate_available_cover" "x2_pool_substrate_characterization"    
#  [5] "x3_pool_variability"                    "x4_sediment_deposition"                 "x5_channel_flow_status"                 "x6_channel_alteration"                 
#  [9] "x7_channel_sinuosity"                   "x8_lb_stability"                        "x8_rb_stability"                        "x9_lb_vegetative_protection"           
# [13] "x9_rb_vegetative_protection"            "x10_lb_riparian_zone_width"             "x10_rb_riparian_zone_width"             "comments"                              

habitat.2016.2017 <- read_excel(fielddata.2016.2017.path, sheet = "Habitat") %>% 
  janitor::clean_names() %>% 
  rename_all(tolower) %>% 
  select(-c(basin, location, rivmile, bas_loc_rm, gradient)) %>% 
  mutate(site_id_field_fixed = str_replace(site_id_field_fixed, "07-HARB-0.5_b", "07-HARB-0.5")) %>% 
  rename(site_id = site_id_field_fixed,
         sample_date = coll_date_habitat,
         epifaunal_cover = epi_cover,
         embeddedness_pooling = embedd_pool,
         velocity_depth_regime = vel_dep_regime,
         sediment_deposition = sed_dep,
         channel_alteration = chan_alt,
         riffle_bend_frequency = rif_bnd_freq,
         left_bank_stability = l_bnk_stab,
         right_bank_stability = r_bnk_stab,
         left_bank_veg = l_bnk_veg,
         right_bank_veg = r_bnk_veg,
         left_bank_veg_zone = l_bnk_zone,
         right_bank_veg_zone = r_bnk_zone,
         habitat_comment = notes)
 

# > names(habitat.2016.2017)
#  [1] "basin"               "location"            "rivmile"             "bas_loc_rm"          "site_id_field_fixed" "coll_date_habitat"   "biosample"          
#  [8] "gradient"            "epi_cover"           "embedd_pool"         "vel_dep_regime"      "sed_dep"             "flow_status"         "chan_alt"           
# [15] "rif_bnd_freq"        "l_bnk_stab"          "r_bnk_stab"          "l_bnk_veg"           "r_bnk_veg"           "l_bnk_zone"          "r_bnk_zone"         
# [22] "habitat_total_score" "notes"  


habitat.bind <- bind_rows(habitat.2015, habitat.2016.2017) 
  # select(sort(current_vars()))

```

Misc ancillary data: Load in and bind as needed.
```{r eval = FALSE}
# Add these into master table (join to 2015 only). Or load into 2015 field data in above.
toc.2015 <- read_excel(fielddata.2015.path, sheet = "TOC")  
grainsz.2015 <- read_excel(fielddata.2015.path, sheet = "GrainSize")

# Need to capture these for 2018-2019 anyway; will run again for all.
ecoregion.2015 <- read_excel(fielddata.2015.path, sheet = "Ecoregion")
ecoregion.2016.2017 <- read_excel(fielddata.2016.2017.path, sheet = "LG2016-17_ecoregion_join") 

# Some StreamStats data was pulled into master tables:
# streamstats.2015 <- read_excel(fielddata.2015.path, sheet = "StreamStats_basins") # Doesn't include all variables from original table but maybe the important ones.
# streamstats.2016.2017 <- read_excel(fielddata.2016.2017.path, sheet = "LG data + sites, LU, SteamStats") %>%
#   select(1,86:90)

# Loading full raw StreamStats output tables:
streamstats.2015 <- read_csv(file.path(root.dir, "data", "compile_2015-2019", "field_data", "StreamStats_table_attributes_2015_96basins_FINAL_1-14-19.csv"))
streamstats.2016.2017 <- read_csv(file.path(root.dir, "data", "compile_2015-2019", "field_data", "StreamStats_table_attributes_2016-17_basins_1-31-19_2.csv"))
streamstats.bind <- bind_rows(streamstats.2015, streamstats.2016.2017)
# NOTE: 	07-LITR-0.2 was run separately (see StreamStats_singlebasin_output_07-LITR-0.2.csv)

rm(field.LU.2015, field.LU.2016.2017, fielddata.2015.path, fielddata.2016.2017.path, habitat.2015, habitat.2016.2017, streamstats.2015, streamstats.2016.2017)

```
Read in bug data (MACRO SP SAMPLE INFO HIST) to get list of 2018-2019 events for filtering field data tables.
```{r}
db.mssih.events <- read_csv(file.path(itstable.path, "20201029_S_MACRO_SPECIES_SAMP_INF_HIST.csv"))  %>% 
  rename_all(tolower) %>% 
  setNames(., sub("mssih_", "", names(.))) %>% 
  rename(site_id = event_smas_history_id,
         sample_date = event_smas_sample_date) %>% 
  mutate(sample_date = as.Date(sample_date, format = "%m/%d/%Y"),
         year = format(sample_date, format = "%Y")) %>% 
  filter(year %in% c("2018", "2019"),
         biosample_collect_method %in% "Low-Gradient",
         # biosample_collect_method_num == 9 (filter result matches above)
         replicate == 1
         ) %>% 
  mutate(linked_id_validator = substr(linked_id_validator,1,nchar(linked_id_validator)-4)) %>% 
  rename(event_id = linked_id_validator) %>% 
  select(event_id) %>% 
  # Fix bad bug sample date
  mutate(event_id = str_replace(event_id, "04-WCONI-0.6_20190722", "04-WCONI-0.6_20190723"))

```


New approach: Join ITS tables and add in 2015-2017 data

Import and trim down SAMPLE EVENT INFO and MACRO FIELD tables to needed fields then join.
```{r}
db.SEI <- read.csv(file.path(itstable.path, "20201019_S_SAMPLE_EVENT_INFO_all_fields.csv"), stringsAsFactors = FALSE) %>% 
  rename_all(tolower) %>% 
  setNames(., sub("seih_", "", names(.))) %>% 
  rename(site_id = event_smas_history_id,
         sample_date = event_smas_sample_date,
         event_id = event_smas_id,
         arrivetime = smas_arrivetime,
         cur_spd = speed_of_stream_current,
         macrophyte_pct = mfdh_macrophytes) %>%
  filter(event_id %in% db.mssih.events$event_id) %>% 
  mutate(sample_date = as.POSIXct(sample_date, format = "%m/%d/%Y"))

db.SEI.select <- db.SEI %>% 
  select(event_id, site_id, basin, sample_date, arrivetime, crew, depth, width, cur_spd, algae_diaoms_presen, thickness_diatoms_present, macrophyte_pct, canopy, embeddedness, rock, rubble, gravel, sand, silt, clay, land_use, general_comment)

db.macrofield <- read.csv(file.path(itstable.path, "20201023_S_MACRO_FIELD_DATA_HISTORY_all_fields.csv"), stringsAsFactors = FALSE) %>% 
  rename_all(tolower) %>% 
  setNames(., sub("mfdh_", "", names(.))) %>%
  rename(event_id = event_smas_id) %>%
  filter(event_id %in% db.mssih.events$event_id) %>% 
  # filtering by the criteria below yields 5 more surveys than bugs present. Look into missing bugs later.
  # mutate(sample_date = as.Date(sample_date, format = "%m/%d/%Y"),
  #        year = format(sample_date, format = "%Y")) %>%
  # filter(year %in% c("2018", "2019"),
  #        biosample_collect_method %in% "Low-Gradient") %>%
  rename(ephem = ephemeroptera,
         plecop = plecoptera,
         trich = trichoptera,
         coleop = coleoptera,
         megalop = megaloptera,
         chir = chironomidae,
         simul = simuliidae,
         decapod = decapods,
         gammar = gammaridae,
         mollusk = mollusca,
         oligo = oligochaeta) %>% 
  # The following field traige columns would not bind; not needed anyway.
  select(-c("ephem", "plecop", "trich", "coleop", "megalop", "odonata", "chir", "simul", "decapod", "gammar", "mollusk", "oligo", "tipulidae", "others", "other"))

db.macrofield.select <- db.macrofield %>% 
  select(-c(row_index, instance_id, event_smas_history_id, basin, location, rivermile, event_smas_sample_date, biosample_collect_method, biosamp_col_method_note))

SEI.MACRO.join <- left_join(db.SEI.select, db.macrofield.select, by = "event_id")

# setdiff(SEI.MACRO.join, field.LU.bind)
# Output:
# Error: not compatible: 
# - Cols in y but not x: `pht_id_dwn`, `pht_id_up`, `hab_present`, `land_use_pct_natural_wetland`, `do_pct_sat`, `cond`, `site_condition_notes`, `cond_test_250`, `site_condition_v1`, `ph`, `temp`, `riffle_present`, `site_condition_v2`, `ref_field_yn`, `salinity`, `photo_2`, `microtox`, `chl_a`, `do_conc`, `photo`. 
# - Cols in x but not y: `macro_comment`, `event_id`. 

#Fields checked; okay.

```

Add in HABAITAT table to above join
```{r}
db.habitat <- read.csv(file.path(itstable.path, "20201020_S_HABITAT_FIELD_DATA_HISTORY_all_fields.csv"), stringsAsFactors = FALSE) %>% 
  rename_all(tolower) %>% 
  setNames(., sub("hfdh_", "", names(.))) %>% 
  rename(site_id = event_smas_history_id,
         sample_date = event_smas_sample_date,
         event_id = event_smas_id) %>% 
  filter(event_id %in% db.mssih.events$event_id) 

db.habitat.select <- db.habitat %>% 
  select(-c(row_index, site_id, basin, location, rivermile, sample_date, instance_id))

SEI.MACRO.HABITAT.join <- left_join(SEI.MACRO.join, db.habitat.select, by = "event_id")
```

Transform IN-SITU data from long to wide and join to above
```{r}
db.insitu <- read.csv(file.path(itstable.path, "20201030_S_IN_SITU_WATER_CHEM_all_fields.csv"), stringsAsFactors = FALSE) %>% 
  rename_all(tolower) %>% 
  setNames(., sub("iswc_", "", names(.))) %>% 
  rename(site_id = event_smas_history_id,
         event_id = event_smas_id,
         sample_date = event_smas_sample_date,
         pcode = chem_parameter_pcode_valid) %>% 
  filter(event_id %in% db.mssih.events$event_id) %>% 
  select(event_id, pcode, result) %>% 
  mutate(pcode = as.character(pcode),
         pcode = case_when(
    pcode == 32 ~ "chl_rfu",
    pcode == 33  ~ "chl_ugL",
    pcode == 61 ~ "do_conc",
    pcode == 62 ~ "do_pct_sat",
    pcode == 109 ~ "ph",
    pcode == 125 ~ "pc_rfu",
    pcode == 126 ~ "pc_ugl",
    pcode == 138 ~ "spcond",
    pcode == 142 ~ "temp",
    pcode == 170 ~ "salinity",
    TRUE ~ pcode
    )
  )

db.insitu.wide <- db.insitu %>% 
  group_by(event_id, pcode) %>% 
  summarise(result = sum(result)) %>% 
  spread(pcode, result, fill = NA) %>% 
  select(event_id, temp, spcond, everything(), -c(chl_rfu, chl_ugL, pc_rfu, pc_ugl))

SEI.MACRO.HABITAT.INSITU <- left_join(SEI.MACRO.HABITAT.join, db.insitu.wide, by = "event_id")

```

Join 2015-2017 field and habitat, bind with master field 2018-2019 table
```{r}
# First join together LG field and habitat data

lg.old.join <- left_join(field.LU.bind, habitat.bind, by = c("site_id", "sample_date"))

# setdiff(SEI.MACRO.HABITAT.INSITU, lg.old.join)

master.field <- bind_rows(SEI.MACRO.HABITAT.INSITU, lg.old.join)

```

Further field data cleaning and export
```{r}

master.field.clean <- master.field %>% 
  select(-c(site_condition_v1, site_condition_v2, site_condition_notes, habitat_total_score)) %>% 
  mutate(bottom_type = tolower(bottom_type),
         bottom_type = str_replace_all(bottom_type, " ", "_"),
         water_clarity_driver = tolower(water_clarity_driver),
         water_clarity_driver = str_replace_all(water_clarity_driver, " ", "_"),
         water_clarity_driver = str_replace_all(water_clarity_driver, "turbidity", "turb"))

# Export to two directories
write.table(master.field.clean, file.path(root.dir, "data", "compile_2015-2019", "field_data", "final_cleaned", "2015-2019_lowgrad_fielddata_MASTER_20201203.csv"),sep=",", row.names = FALSE)
write.table(master.field.clean, file.path(teams.datadir, "field-sites", "2015-2019_lowgrad_fielddata_MASTER_20201203.csv"),sep=",", row.names = FALSE)

```

Sites table
```{r}
# Only 163 unique sites present.
# Identify which sites were sampled twice and which are new in 2018-2019
master.field.sitecount <- master.field %>% 
  select(site_id, sample_date, basin) %>% 
  group_by(site_id) %>% 
  mutate(sampled_count = n()) %>% 
  mutate(year = format(sample_date, format = "%Y")) %>% 
  mutate(newsite.2018.2019 = ifelse(year %in% c("2018", "2019") & sampled_count!=2, 1, 0)) %>% 
  select(-c(sample_date, basin, year)) %>% 
  distinct()

sites.lg <- read_excel(file.path(itstable.path, "20201030_S_SITE.xlsx")) %>% 
  rename_all(tolower) %>% 
  setNames(., sub("site_", "", names(.))) %>%
  rename(site_id = history_id) %>%
  select(site_id, latitude, longitude, basin, name, desc, huc8, water_qlty_standard, habitat_type, site_condition) %>% 
  filter(site_id %in% master.field$site_id) %>% 
  left_join(master.field.sitecount, by = "site_id")

# Subset to sites new in 2018-2019 for performing land use analysis
sites.lg.new.in.2018.2019 <- sites.lg %>% 
  filter(newsite.2018.2019 == 1)

# List sites that are IDd as high gradient in the sites table
sites.lg.HG <- sites.lg %>% 
  filter(grepl("*High", habitat_type))




### Add in elev, slope, other info from StreamStats ####
# Old files in C:\Data\LGSS_scripting\lgss\data\compile_2015-2019\steamstats_basinstats 
#          and ...\field_data
# Export 2018-2019 stats from shapefile in ArcGIS Pro (already did?)



# write.table(sites.lg, file.path(teams.datadir, "sites.lowgradient_20201104.csv"),sep=",", row.names = FALSE)
# write.table(sites.lg.new.in.2018.2019, file.path(teams.datadir, "field-sites", "sites.lowgradient.newin20182019_20201104.csv"),sep=",", row.names = FALSE)


```
