---
title: "compile_2016-17"
author: "Gavin Lemley"
date: "January 2, 2019"
output: html_document
---

### For abiotic data ###
Find the R-project root directory to simplify file path designations.
```{r}
root.dir <- rprojroot::find_root("lgss.Rproj")
```

Load in data tables and create BAS_LOC_RM column in sites table.
```{r}

field_data <- read.csv(file.path(root.dir, "data/compile_2016-17", "LGSS_field_2016_2017.csv"))
sites_tbl <- read.csv(file.path(root.dir, "data/compile_2016-17", "Sites_DiUpdated_2018-12-17.csv"))

#Format fields to add leading, and keep trailing zeros.
sites_tbl$BASIN <- formatC(sites_tbl$BASIN, width = 2, format = "d", flag = "0")
sites_tbl$RIVMILE <- formatC( sites_tbl$RIVMILE, format='f', digits=1 )

#Create BAS_LOC_RM field
sites_tbl$BAS_LOC_RM <- paste0(sites_tbl$BASIN,"-",sites_tbl$LOCATION,"-",sites_tbl$RIVMILE)

```

Merge tables to associate site info with field data.
```{r}
FSmerge <- merge(field_data,sites_tbl,by="BAS_LOC_RM", all=FALSE)

```

```{r}
write.table(FSmerge, file.path(root.dir, "data/compile_2016-17/LGSS_16-17_field-Sites_merge.csv"),sep=",", row.names = FALSE)

```

### For 2016-17 bug data (long format to matrix for PRIMER import) ###

Specify the file path to the XLS file.
```{r}
root.dir <- rprojroot::find_root("lgss.Rproj")

bugs.path <- file.path(root.dir,
                            "data",
                            "compile_2016-17",
                            "LGSS_BUGS_2016_2017_IDSs_fixed.xlsx")
```

Identify all of the sheets in the XLS file.
```{r}
sheet.vec <- readxl::excel_sheets(bugs.path)
```

Read in 2016 and 2017 bugs sheets (long format), bind, and create site ID field
```{r}
bugs2016.df <- readxl::read_excel(bugs.path, sheet = "LGSS 2016")
bugs2017.df <- readxl::read_excel(bugs.path, sheet = "LGSS 2017")

bugs201617.df <-rbind(bugs2016.df,bugs2017.df)

#Format fields to add leading, and keep trailing zeros.
bugs201617.df$BASIN <- as.numeric(bugs201617.df$BASIN)
bugs201617.df$BASIN <- formatC(bugs201617.df$BASIN, width = 2, format = "d", flag = "0")

bugs201617.df$RIVMILE <- as.numeric(bugs201617.df$RIVMILE)
bugs201617.df$RIVMILE <- formatC( bugs201617.df$RIVMILE, format='f', digits=1 )

#Create BAS_LOC_RM field
bugs201617.df$BAS_LOC_RM <- paste0(bugs201617.df$BASIN,"-",bugs201617.df$LOCATION,"-",bugs201617.df$RIVMILE)

# Create year field for troubleshooting missing data in ITS tables
bugs201617.df_TEMP <- bugs201617.df %>% 
  mutate(YEAR = format(as.Date(COLL_DATE, format="%m/%d/%Y"),"%Y")) %>% 
  select(BAS_LOC_RM, YEAR) %>% 
  distinct() %>% 
  select(YEAR)

```

Transform bugs from long to matrix
```{r}
library(tidyverse)
bugmatrix1617.df <- bugs201617.df %>% 
  dplyr::select(BAS_LOC_RM, MACRO_GENSPECIES, INDIV) %>% 
  group_by(BAS_LOC_RM, MACRO_GENSPECIES) %>% 
  summarise(INDIV = sum(INDIV)) %>% 
  # mutate(MACRO_GENSPECIES = str_replace_all(MACRO_GENSPECIES, " ", "_")) %>% 
  spread(MACRO_GENSPECIES, INDIV, fill = 0)

#Rename repeat site (already sampled in 2015)
bugmatrix1617.df$BAS_LOC_RM <- gsub("07-HARB-0.5", "07-HARB-0.5_b",bugmatrix1617.df$BAS_LOC_RM)

#Count total to check if >200
# bugmatrix1617.df$total <- rowSums(bugmatrix1617.df[,2:280])

#Replace spaces in column names to match 2015 data
# names(bugmatrix1617.df) <- gsub(x = names(bugmatrix1617.df), pattern = "\\ ", replacement = ".")
  
# write.table(bugmatrix1617.df, file.path(root.dir, "data/LG_2016-17_BUGS.csv"),sep=",", row.names = FALSE)


```

Bind 2015 bugs with 2016-17
```{r}
bugmatrix2015.df <- read.csv(file.path(root.dir, "data", "2015_bugs.csv"), stringsAsFactors = FALSE, check.names = FALSE)
#Replace NAs with 0s
bugmatrix2015.df[is.na(bugmatrix2015.df)] <- 0
#Rename repeat site (also sampled in 2016-17 dataset)
bugmatrix2015.df$BAS_LOC_RM <- gsub("07-HARB-0.5", "07-HARB-0.5_a",bugmatrix2015.df$BAS_LOC_RM)

LG_all_bugs.df <- dplyr::bind_rows(bugmatrix2015.df, bugmatrix1617.df)
LG_all_bugs.df[is.na(LG_all_bugs.df)] <- 0
LG_all_bugs.df <- LG_all_bugs.df[order(LG_all_bugs.df$BAS_LOC_RM),] 

write.table(LG_all_bugs.df, file.path(root.dir, "data/LG_ALL_BUGS.csv"),sep=",", row.names = FALSE)

```

Invesitgating why bind results in different number of taxa when converting spaces to decimals (line 98)... 
```{r}
# bugmatrix2015.df.chkFALSE <- read.csv(file.path(root.dir, "data", "2015_bugs.csv"), check.names = FALSE) %>% 
#   dplyr::bind_rows(bugmatrix1617.df)
# names(bugmatrix2015.df.chkFALSE) <- str_replace_all(names(bugmatrix2015.df.chkFALSE), " ", "_")
# 
# bugmatrix2015.df.chkTRUE <- read.csv(file.path(root.dir, "data", "2015_bugs.csv"), check.names = TRUE) %>% 
#     dplyr::bind_rows(bugmatrix1617.df)
# 
# 
# test <- data.frame(name = names(LG_all_bugs.df))
# test2 <- data.frame(name = names(bugmatrix2015.df))
# 
# anti.df <- anti_join(test, test2)
# 
# name!names(bugmatrix2015.df) %in% names(bugmatrix1617.df)
```

