---
title: "Untitled"
author: "Zachary M. Smith"
date: "December 21, 2018"
updated: "12/26/18 by Gavin Lemley""
output: html_document
---
`r format(Sys.Date(), "%m-%d-%Y")`

# Overview

This document was created to import the land cover data calculated by C. Stoll and G. Lemley for the LGSS project.

# Preparation

Import libs and find root
```{r}
library(tidyverse)
library(readxl)

root.dir <- rprojroot::find_root("lgss.Rproj")

```

Specify the file path to the XLS file.
```{r}
landcover.15.path <- file.path(root.dir,
                            "data",
                            "compile_2015-2019",
                            "landcover",
                            "Batch_Land_Cover_Output_12212018_2015basins.xls")

landcover.16.17.path <- file.path(root.dir,
                            "data",
                            "compile_2015-2019",
                            "landcover",
                            "Batch_Land_Cover_Output_01312019_2016-17basins.xls")

landcover.18.19.path <- file.path(root.dir,
                            "data",
                            "compile_2015-2019",
                            "landcover",
                            "Batch_Land_Cover_Output_11182020_2018-19basins.xls")
```

Identify all of the sheets in the XLS file.
```{r}
sheet.vec.15 <- readxl::excel_sheets(landcover.15.path)
sheet.vec.16.17 <- readxl::excel_sheets(landcover.16.17.path)
sheet.vec.18.19 <- readxl::excel_sheets(landcover.18.19.path)

```

Import all of the sheets in the XLS file and bind them into a single data frame. The sheet name is used to generate the `site_id` column. The prefix "Basin_" is dropped from the `site_id` column.
```{r}
landcover.15.df <- purrr::map_df(sheet.vec.15, function(sheet.i) {
  readxl::read_xls(landcover.15.path,
                   sheet = sheet.i) %>% 
    mutate(site_id = sheet.i,
           site_id = str_replace(site_id, "Basin_", ""))
})

landcover.16.17.df <- purrr::map_df(sheet.vec.16.17, function(sheet.i) {
  readxl::read_xls(landcover.16.17.path,
                   sheet = sheet.i) %>% 
    mutate(site_id = sheet.i,
           site_id = str_replace(site_id, "Basin_", ""))
})

landcover.18.19.df <- purrr::map_df(sheet.vec.18.19, function(sheet.i) {
  readxl::read_xls(landcover.18.19.path,
                   sheet = sheet.i) %>% 
    mutate(site_id = sheet.i,
           site_id = sub('.', '', site_id))
})

landcover.df <- bind_rows(landcover.15.df, landcover.16.17.df, landcover.18.19.df)

# total = 167. 3 sites missing (should be 169 but includes 1 dup site to check reproducability)

```

```{r}
landcover.df <- landcover.df %>% 
  group_by(site_id) %>% 
  mutate(total_acres = sum(Area_acre)) %>% 
  ungroup() %>% 
  mutate(percent_cover = Area_acre / total_acres * 100)

```

```{r}
NLCD_key <- read.csv(file.path(root.dir, "data", "compile_2015-2019", "landcover", "NLCD_attribute_key_condensed_LGSS_v2.csv"))

landcover.df <- merge(landcover.df,NLCD_key, by = "Value", all.x = TRUE)

```

Old method
```{r eval = FALSE}
sitetype.df <- landcover.df %>% 
  filter(Condensed_class %in% c("Natural", "Wetland")) %>% 
  group_by(site_id) %>% 
  summarize(pct = sum(percent_cover)) %>% 
  mutate(site_type = ifelse(pct>=75,"ref","test"))

# write.table(sitetype.df, file.path(root.dir, "data/compile_2016-17/LGSS_16-17_landuse.csv"),sep=",", row.names = FALSE)


```
New method including all "condensed categories" with new NLCD_key v2
```{r}
landcover.nat <- landcover.df %>% 
  filter(Condensed_class %in% "Natural") %>% 
  group_by(site_id) %>% 
  summarize(landcover_nat_pct = sum(percent_cover))

landcover.dev <- landcover.df %>% 
  filter(Condensed_class %in% "Developed") %>% 
  group_by(site_id) %>% 
  summarize(landcover_dev_pct = sum(percent_cover))

landcover.water <- landcover.df %>% 
  filter(Condensed_class %in% "Water") %>% 
  group_by(site_id) %>% 
  summarize(landcover_water_pct = sum(percent_cover))

landcover.ag <- landcover.df %>% 
  filter(Condensed_class %in% "Agricultural") %>% 
  group_by(site_id) %>% 
  summarize(landcover_ag_pct = sum(percent_cover))

landcover.all <- left_join(landcover.nat, landcover.dev, by = "site_id") %>% 
  left_join(., landcover.ag, by = "site_id") %>% 
  left_join(., landcover.water, by = "site_id") %>% 
  # Add all to verify 100%
  mutate(total_pct = select(., landcover_nat_pct, landcover_dev_pct, landcover_water_pct, landcover_ag_pct) 
         %>% rowSums(na.rm = TRUE)) %>% 
  select(-total_pct) %>% 
  # Remove reproducability site
  filter(!(site_id %in% "08_BENE_0_2_repro"))

```

```{r}
write.table(landcover.all, file.path(root.dir, "data/compile_2015-2019/landcover/landcover_final_2015-2019_20201118.csv"),sep=",", row.names = FALSE)

```

